---
layout: post
title: 2016 NCAAF Recruiting Geographic Breakdown
date: 2016-06-04 19:30
category: Sports
tags: data-analysis, ncaaf, data-viz
---

```{r, echo=FALSE}
knitr::opts_chunk$set(fig.path='{{ site.url }}/images/recruiting-geography-', message = FALSE, warning=FALSE)
```

With college football less than 100 days away, here is a look at just how far 250 of the nation's top incoming prospects have to travel to get to their chosen colleges.

The prospect data come courtesy of [Rivals](https://n.rivals.com/prospect_rankings/rivals250/2016), while I obtained school location data from [Wikipedia](https://en.wikipedia.org/wiki/List_of_NCAA_Division_I_FBS_football_programs).

The small blue dots represent hometowns and the larger black stars represent colleges and universities.

R Code is included below while my data wrangling code is posted to my [GitHub](https://github.com/jskaza/jskaza.github.io/tree/source/content/supps/Recruiting/recruits_data.ipynb).

```{r recruits_map, out.width = '100%', dpi = 300}
library(maps)
library(geosphere)

master = read.csv("../data/recruiting-geography/master.csv")
col = rgb(7, 7, 7, 50, maxColorValue = 255)

xlim <- c(-125, -65)
ylim <- c(24, 50)

map("world", col="#f2f2f2", fill=TRUE, bg="white", lwd=0.05, xlim=xlim, ylim=ylim)
for (i in 1:length(master$X)) {
inter = gcIntermediate(c(master[i,]$home_lng, master[i,]$home_lat),
                       c(master[i,]$school_lng, master[i,]$school_lat), n=100, addStartEnd=TRUE)
   
lines(inter, col=col, lwd=1)
}

points(master$home_lng, master$home_lat, col='blue', cex=.25, pch=16)
points(master$school_lng, master$school_lat, col='black', cex=.5, pch=8)
```